Total biases resulting from selection and misclassification errors and according
to given disease prevalence, Se, and Sp are illustrated for disease incidence
and RR in \Cref{fig:incidence_contour,fig:incidence_risk}, respectively.
These figures are contour plots where the lines are curves in the \(x, y\)-plane
along which the function of the two variables on the vertical and horizontal
axes (i.e.\ Se and Sp) has a constant value, i.e.\ a curve joins points of equal
value~\citep{courant_1996}.
The true incidence rate (or RR) is therefore to be found at the upper right
corner of the plot.
For example, in the bottom left panel of \Cref{fig:incidence_contour} the second
line from the bottom is labelled 0.22.
This line shows that, for a 5\% disease prevalence and a true incidence rate of
0.1 case/animal-time unit, an apparent incidence estimate of 0.22 will be
achieved by any combination of Sp and Se on this line (e.g.\ Sp = 0.845, Se =
0.7 or Sp = 0.87 and Se = 1.00).
As an other example, in the upper right panel of this same figure, the first
line at the top is labelled 0.02.
It shows that, for a 5\% disease prevalence and a true incidence rate of 0.01
case/animal-time unit, an apparent incidence estimate of 0.02 is achieved along
this line by any combination of Se and Sp like, for example, a Sp of 1.00 and a
Se of 0.955. The true incidence rate is given at the upper right corner, where
Se and Sp are both 100\%.
Imperfect Se to identify individuals at risk at baseline and imperfect Sp to
identify incident cases led to a mild under-estimation of the observed disease
incidence (Figures~S1 and S2 in Supplementary Material).
From these graphs we could also note that Sp has little effect on selection bias
while Se has little effect on misclassification bias.
Of the two, misclassification bias had a much bigger effect than selection bias.
But overall, the combination of the two biases, at baseline and follow-up,
revealed the importance of a good to excellent Sp relative to Se for the
diagnostic test.
Small divergence from perfect Sp extended quickly to disease incidence
over-estimation as true prevalence increased and true incidence decreased
(\Cref{fig:apparent_incidence_5,fig:apparent_incidence_20,fig:apparent_RR}).
Selection and misclassification biases of a low prevalent and incident disease,
diagnosed with close to perfect Sp, were minimal, reflecting the importance of
choosing a highly specific test to improve identification of animal (or
individual) unit at risk and incident case identification.
The same effect was also observed with RR estimations (Figures~S3 and S4).
Similar results were found with a weaker exposure, RR of 1.5 (Figures~S5 to S8).

%%% Local Variables:
%%% ispell-local-dictionary: "canadian"
%%% mode: latex
%%% TeX-master: t
%%% reftex-default-bibliography: ("./bias.bib")
%%% End:
